parzer
======

```{r echo=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  warning = FALSE,
  message = FALSE
)
```

[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Build Status](https://travis-ci.com/ropenscilabs/parzer.svg?branch=master)](https://travis-ci.com/ropenscilabs/parzer)
[![codecov.io](https://codecov.io/github/ropenscilabs/parzer/coverage.svg?branch=master)](https://codecov.io/github/ropenscilabs/parzer?branch=master)

`parzer` parses messy coordinates

You may get data from a published study or a colleague, and the coordinates
may be in some messy format that you'd like to clean up to e.g., have 
all decimal degree numeric data.

`parzer` API:

```{r echo=FALSE, comment=NA, results='asis'}
cat(paste(" -", paste(sprintf("`%s`", sort(getNamespaceExports("parzer"))), collapse = "\n - ")))
```


## Installation

```{r eval=FALSE}
remotes::install_github("ropenscilabs/parzer")
```

```{r}
library("parzer")
```

## parse

latitudes

```{r warning=TRUE}
parse_lat("45N54.2356")
parse_lat("-45.98739874")
parse_lat("40.123°")
parse_lat("40.123N")
parse_lat("N45 04.25764")

# bad values -> NaN
parse_lat("191.89")

# many inputs
x <- c("40.123°", "40.123N74.123W", "191.89", 12, "N45 04.25764")
parse_lat(x)

# parse_lat("N455698735", "HDDMMmmmmm") # custom formats not ready yet
```

longitudes

```{r warning=TRUE}
parse_lon("45W54.2356")
parse_lon("-45.98739874")
parse_lon("40.123°")
parse_lon("74.123W")
parse_lon("W45 04.25764")

# bad values
parse_lon("181")

# many inputs
x <- c("45W54.2356", "181", 45, 45.234234, "-45.98739874")
parse_lon(x)

# parse_lon("N455698735", "HDDMMmmmmm") # custom formats not ready yet
```

both lat and lon together

```{r}
lats <- c("40.123°", "40.123N", "191.89", 12, "N45 04.25764")
lons <- c("45W54.2356", "181", 45, 45.234234, "-45.98739874")
parse_lat_lon(lats, lons)
```

parse into degree, min, sec parts

```{r warning=FALSE}
parse_parts_lat("45N54.2356")
parse_parts_lon("-74.6411133")
# many inputs
x <- c("40.123°", "40.123N74.123W", "191.89", 12, "N45 04.25764")
parse_parts_lon(x)
```

get hemisphere from lat/lon coords

```{r}
parse_hemisphere("45N54.2356", "74.123E")
parse_hemisphere("40.4183318", "-120")
parse_hemisphere("-40.4183318", "-120")
parse_hemisphere("-40.4183318", "120")
```

get degree, minutes, or seconds separately

```{r}
coords <- c(45.23323, "40:25:6N", "40° 25´ 5.994\" N")
pz_degree(coords)
pz_minute(coords)
pz_second(coords)

coords <- c(15.23323, "40:25:6E", "192° 25´ 5.994\" E")
pz_degree(lon = coords)
pz_minute(lon = coords)
pz_second(lon = coords)
```

add or subtract degrees, minutes, or seconds

```{r}
d(31)
d(31) + m(44)
d(31) - m(44)
d(31) + m(44) + s(59)
d(-121) + m(1) + s(33)
```

## Similar art

- `sp::char2dms`: is most similar to `parzer::parse_lat` and `parzer::parse_lon`. However,
with `sp::char2dms` you have to specify the termination character for each of degree,
minutes and seconds. `parzer` does this for the user.
- `biogeo::dms2dd`: very unlike functions in this package. You must pass separate degrees,
minutes, seconds and direction to `dms2dd`. No exact analog is found in `parzer`, whos
main focus is parsing messy geographic coordinates in strings to a more machine readable
version

## Meta

* Please [report any issues or bugs](https://github.com/ropenscilabs/parzer/issues).
* License: MIT
* Get citation information for `parzer` in R doing `citation(package = 'parzer')`
* Please note that this project is released with a [Contributor Code of Conduct][coc]
By participating in this project you agree to abide by its terms.

[coc]: https://github.com/ropenscilabs/parzer/blob/master/CODE_OF_CONDUCT.md
